Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}

### 2020-09-08 journal paper/content/related_work.tex
Another pass on related work.

\xhdr{Exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information. MAB problems have been studied for many decades by researchers from computer science, operations research, statistics and economics, generating a vast and multi-threaded literature.  The most relevant thread concerns the basic model of regret-minimizating bandits with stochastic rewards and no auxiliary structure (which is the problem faced by each principal in our model), see Appendix~\ref{app:bg} for background. This basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.%
\footnote{There is a superficial similarity, in name only, between this paper and the work on ``dueling bandits" (starting from \citet{Yue-dueling12,Yue-dueling-icml09}).
In the latter, an algorithm chooses two arms in each round, and the algorithm only observes which arm has ``won the duel".}
Dedicated monographs \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} cover the work on regret-minimizing formulations (which mainly comes from computer science), and the classic book
\citep{Gittins-book11} focuses on the Markovian formulations (which predate regret-minimization and were mainly studied in operations research and statistics).
Various connections to economics and game theory are detailed in
books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. Industrial applications are discussed in \citep{DS-arxiv}.

@@

\cite{Ufuk-jeea15} study exploration in a ``research competition". Two firms race towards a big discovery such as a new drug. Each firm solves a bandit problem with two arms, corresponding to safe and risky lines of research. The two algorithms do not interact, directly or indirectly, until one firm makes the discovery and wins the game. In comparison, we study a ``product competition" in which the two firms interact continuously (via the customers' choices), accrue rewards incrementally, and compete for individual customers.

In the line of work on \emph{strategic experimentation} (starting from \citet{Bolton-econometrica99,Keller-econometrica05}, see \citet{Horner-survey16} for a survey), agents explore and learn over time in a shared environment. Thus, we have exploration algorithms which interact with each other strategically, \eg each agent would prefer to free-ride on the exploration done by others. However, the agents do not compete with each other in any meaningful sense.

A long line of work from electrical engineering and computer science, starting from
\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, targets an application to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth. Each radio chooses channels over time using a multi-armed bandit algorithm. This work studies a repeated game between bandit algorithms, and focuses on designing algorithms which work well in this game, under various assumptions on communication, synchronization and collisions.

@@

\xhdr{Competition.} The competition vs. innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{aghion2005competition,Vives-08}. Production costs aside, this literature treats innovation as a priori beneficial for the firm. Our setting is very different, as innovation in exploration algorithms may potentially hurt the firm.

@@

\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers. We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry. In our model the first-mover advantage further provides the incumbent with a ``head start" on data collection relative to the late arrivers. Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{de2020data}. For an extensive overview of the other channels through which first-mover advantages can lead to a competitive advantage, see \cite{kerin1992first}.

@@

The ``dueling algorithms" framework of \citet{DuelingAlgs-stoc11} studies competition between two principals, each running an algorithm for the same problem. However, they consider algorithms for offline / full input scenarios, and posit binary (win/lose) payoffs for the principals. Similarly, \citet{ben2017best, ben2019regression} study competition between offline learning algorithms. Whereas we focus on bandit algorithms and competition for users.

>>>>

\xhdr{Exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information. MAB problems have been studied for many decades by researchers from computer science, operations research, statistics and economics, generating a vast and multi-threaded literature.  The most relevant thread concerns the basic model of regret-minimizating bandits with stochastic rewards and no auxiliary structure (which is the problem faced by each principal in our model), see Appendix~\ref{app:bg} for background. This basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.%
\footnote{There is a superficial similarity, in name only, between this paper and the work on ``dueling bandits" \citep[starting from][]{Yue-dueling12,Yue-dueling-icml09}.
In this work, there is only one bandit algorithm which chooses two arms in each round, and the only observes which arm has ``won the duel".}
Dedicated monographs \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} cover the work on regret-minimizing formulations (which mainly comes from computer science). The classic book
\citep{Gittins-book11} focuses on the Markovian formulations (which predate regret-minimization and were mainly studied in operations research and statistics).
Various connections to economics and game theory are detailed in
books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. Industrial applications are discussed in \citep{DS-arxiv}.

@@

\vspace{2mm}

\asmargincomment{It serves us to discuss this work here, IMO, and we have an excuse.}
In the line of work on \emph{strategic experimentation} (starting from \citet{Bolton-econometrica99,Keller-econometrica05}, see \citet{Horner-survey16} for a survey), agents explore and learn over time in a shared environment. Thus, we have exploration algorithms which interact with each other strategically, \eg each agent would prefer to free-ride on the exploration done by others. However, the agents do not compete with each other in any meaningful sense.

Several papers study competition between two principals who run algorithms but do not interact, directly or indirectly, until the very end of the game. \cite{Ufuk-jeea15} consider a ``research competition" between two firms racing towards a big discovery. 
\asmargincomment{This way, this paper seems not that relevant, doesn't it?}
Each firm deploys a bandit algorithm with two arms, corresponding to safe and risky lines of research. The firms do not interact untl one of them makes the discovery and wins the game. In the ``dueling algorithms" framework of \citet{DuelingAlgs-stoc11}, each principals runs an algorithm for the same problem. All inputs are observable at once, and principals' payoffs are binary (win/lose). \citet{ben2017best, ben2019regression} study competition between ``offline" machine learning algorithms. In comparison, we study a ``product competition" in which the two firms interact continuously (via the customers' choices), accrue rewards incrementally, and compete for individual customers.

\asmargincomment{This happens to be about exploration \& competition ...}
A long line of work from electrical engineering and computer science, starting from
\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, targets an application to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth. Each radio chooses channels over time using a multi-armed bandit algorithm. This work studies a repeated game between bandit algorithms, and focuses on designing algorithms which work well in this game, under various assumptions on communication, synchronization and collisions.

@@

\xhdr{Competition in economics.} The competition vs. innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{aghion2005competition,Vives-08}. Production costs aside, this literature treats innovation as a priori beneficial for the firm. Our setting is very different, as innovation in exploration algorithms may potentially hurt the firm.

@@

\asedit{\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers; we find a similar effect with ``reputation advantage". The role of data as a barrier to entry in online markets has been noted in \cite{de2020data}; we find a similar effect with ``data advantage". \citet{kerin1992first} overview the other channels through which first-mover advantage can lead to a competitive advantage.}

%\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers. We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry. In our model the first-mover advantage further provides the incumbent with a ``head start" on data collection relative to the late arrivers. Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{de2020data}. For an extensive overview of the other channels through which first-mover advantages can lead to a competitive advantage, see \cite{kerin1992first}.

@@
